---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in the FAS Litters Data

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## `select`

Choose some columns but not others.

```{r}
select(litters_df, group, litter_number)
```

Choose some columns in sequence
```{r}
select(litters_df, gd0_weight:gd_of_birth)
```

Get rid of a(some) column(s)
```{r}
select(litters_df, -litter_number, -group)
```

Renaming columns
```{r}
select(litters_df, GROUP = group, LITTer_NUmBeR = litter_number)
```

Renaming some columns while keeping all columns 
```{r}
rename(litters_df, GROUP = group, LITTer_NUmBeR = litter_number)
```

Select Helpers (ask **?select_helpers** in console for more information)
```{r}
select(litters_df, starts_with("gd"))
```

Move (select) one column to the beginning
```{r}
select(litters_df, litter_number, everything())
```

Another function which does the exact same thing as above
```{r}
relocate(litters_df, litter_number)
```


## `filter`

select rows based on whether some observations is true
```{r}
filter(litters_df, gd0_weight < 22)
```
```{r}
filter(litters_df, gd0_weight >= 22)
```

When we ask whether a variable equals to what value, we use **two** equal signs
```{r}
filter(litters_df, gd_of_birth == 20)
```

When we want the variable with values which are not equal to a particular number
```{r}
filter(litters_df, !(gd_of_birth) == 20)
```
Another way to do things above
```{r}
filter(litters_df, gd_of_birth != 20)
```

When we want multiple statements to be true at the same time
```{r}
filter(litters_df, gd_of_birth == 20, gd0_weight >= 22)
```

When we want to extract specific values
```{r}
filter(litters_df, group == "Con7")
```

When we combine things together
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```


## `mutate`

Add a variable
```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)
```

Create/edit multiple variables at the same time
```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```


## `arrange`

```{r}
arrange(litters_df, pups_born_alive)
```
If you want to arrange multi-variables, the variable you put first will be arranged firstly.
```{r}
arrange(litters_df, pups_born_alive, gd0_weight)
```


## `%>%`

```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight)
```

USE THE PIPE OPERATOR INSTEAD

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>%
  select(-pups_survive) %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
```












